Hierarchical Dispatching and Overriding

### 2017-10-04 NewVersion/sections/Formalization.tex
refine s3

\section{Formalization}~\label{sec:formalization}
In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}), based on
Featherweight Java~\cite{Igarashi01FJ}. The syntax, typing rules,
small-step semantics and type soundness proofs are presented.

\vspace{-2ex}
\subsection{Syntax}
The abstract syntax of \MIM{} interface declarations, method declarations, and expressions is given at Figure~\ref{fig:syntax}. The multiple
inheritance feature of \MIM{} is inspired by Java 8 interfaces, which support
method implementations via default methods. This feature is 
closely related to \emph{traits}. To demonstrate how
unintentional method conflicts are untangled in \MIM{}, we only focus on a small subset of the interface model. For example, all the methods declared
in an interface are either default methods or abstract methods. Default methods provide default implementations for methods, and abstract methods have no method body, which requires future implementation. 
%From this point we can view that we
%are actually modelling a class model that supports multiple
%inheritance. 
Object creation is directly supported via $\new I$. 
Fields and primitive types are not modelled in \MIM{}.

@@

The method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method is the single statement $ \kwreturn \; e;$. We also allow
defining abstract methods with $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a normal method but without a method body. Notably, we have introduced the
\kwoverride{} keyword to override an old implementation of the
method. If the interface that it overrides is exactly the enclosing
interface, then such a method is seen as ``originally defined''; otherwise
it is hierarchical overriding.
For simplicity, overloading is not modelled for methods, which
implies we can uniquely identify a method by its name.

@@

Details of type-checking rules are displayed at the bottom of Figure~\ref{fig:typingrules}, including expression
typing, well-formedness of methods and interfaces. As a convention, an environment
$\Gamma$ is maintained to store the types of variables, together with
the ``this'' type, namely
the enclosing type. The three rules for method invocation, \textsc{(T-Invk)}, \textsc{(T-PathInvk)} and \textsc{(T-SuperInvk)}
are very similar, in the sense that they all check the type of the specific method, by using
an auxiliary function \mtype. \mtype{} is the function for looking up method types, which we will
illustrate later in Section~\ref{subsec:auxdefs}. After the method
type is obtained, they all check that the arguments and the receiver
have compatible types. Additionally, \textsc{(T-PathInvk)} requires the receiver to be the subtype of the specified
path type, and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

	\framebox{$ \judgeewf \Gamma {e:I} $} \hspace{.5in}
	\tvar \\
	\tinvk \\
	\tpathinvk \\
	% \tsuperinvk \\
	% \tstaticinvk  \\
	\tnew \\
	\tmethod \\
	\tabsmethod \\
	\tintf \\
	\tanno
\end{mathpar}
\caption{Typing Rules of \name{}.}
\label{fig:typingrules}
\saveSpaceFig
\caption{Typing and subtyping of \name{}.}
\end{figure*}

@@

Generally, one can understand
$\mbody(m, I, J)$ in a way that it finds the most specific body of
method $m$, when the receiver has dynamic type $I$ and static type
$J$.  Notice, for example in \textsc{(S-Invk)}, that when
``$\new{I}$'' replaces ``\lstinline|this|'', its static type should be
the interface to which $m$ is dispatched. Therefore we also keep the
interface type in the definition of \mbody. This point is consistent
with the last paragraph in Section~\ref{subsec:problem2}.

On the other hand, there is a relationship between path invocation and
the regular method invocation, and it can be observed from the
similarity between their semantic rules in
Figure~\ref{fig:smallstep}. For any \lstinline|e.I::m()|, the result
of evaluation remains unchanged if we set the static type of
\lstinline|e| to be $I$. This can be done by an implicit cast, that
is, we can define a function with one parameter type $I$, then
\lstinline|e| is passed to that function and directly returned. This
is equivalent to writing explicit casts like \lstinline|((I) e).m()|
in languages like Java.

@@

\subsection{Properties}
Previously the definitions of our model are given, now we should proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be a subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the number of
elements.

>>>>

\section{Formalization}~\label{sec:formalization}
In this section, we present the formal model \MIM{} (\emph{\textbf{F}eatherweight \textbf{H}ierarchical \textbf{J}ava}), based on
Featherweight Java~\cite{Igarashi01FJ}. The syntax, typing rules,
small-step semantics and theorems for type soundness are presented.

\vspace{-2ex}
\subsection{Syntax}
The abstract syntax of \MIM{} interface declarations, method declarations, and expressions is given at Figure~\ref{fig:syntax}. The multiple
inheritance feature of \MIM{} is inspired by Java 8 interfaces, which support
method implementations via default methods. This feature is 
closely related to \emph{traits}. To demonstrate how
unintentional method conflicts are untangled in \MIM{}, we only focus on a small subset of the interface model. For example, all the methods declared
in an interface are either default methods or abstract methods. Default methods provide default implementations for methods. Abstract methods do not
have method body, and in fact we are not introducing a different kind of methods, but they are just treated as normal methods that return empty. Abstract methods can be overridden with future implementations.

%From this point we can view that we
%are actually modelling a class model that supports multiple
%inheritance. 
Object creation is represented by $\new I$. If any unimplemented abstract methods are defined in interface $I$ (or inherited from parents), then object creation fails. Fields and primitive types are not modelled in \MIM{}.

@@

The method declaration $\method{I}{m}{I_x}{x}{J}{e}$ introduces a
method named $m$ with result type $I$, parameters $\overline{x}$ of
type $\overline{I_x}$ and the overriding target $J$. The body of the
method is the single statement $ \kwreturn \; e;$. We also allow
defining abstract methods with $\absmethod{I}{m}{I_x}{x}{J}$, which is
similar to a normal method but without a method body. Notably, we have introduced the
\kwoverride{} keyword to override an old implementation of the
method. If the interface that it overrides is exactly the enclosing
interface, then such a method is seen as ``originally defined''; otherwise
it is hierarchical overriding.
For simplicity, overloading is not modelled for methods, which
implies we can uniquely identify a method by its name.\yanlin{the terminology of original methods should have been proposed in Overview.}

@@

Details of type-checking rules are displayed at the bottom of Figure~\ref{fig:typingrules}, including expression
typing, well-formedness of methods and interfaces. As a convention, an environment
$\Gamma$ is maintained to store the types of variables, together with
the \kwthis type, namely
the enclosing type. 
% The three rules for method invocation, \textsc{(T-Invk)}, \textsc{(T-PathInvk)} and \textsc{(T-SuperInvk)}
% are very similar, in the sense that they all check the type of the specific method, by using
% an auxiliary function \mtype. \mtype{} is the function for looking up method types, which we will
% illustrate later in Section~\ref{subsec:auxdefs}. After the method
% type is obtained, they all check that the arguments and the receiver
% have compatible types. Additionally, \textsc{(T-PathInvk)} requires the receiver to be the subtype of the specified
% path type, and \textsc{(T-SuperInvk)} checks if the enclosing type directly extends the specified super type.

@@

	\framebox{$ \judgeewf \Gamma {e:I} $} \hspace{.5in}
	\tvar \\
	\tinvk \\
	\tpathinvk \\
	% \tsuperinvk \\
	% \tstaticinvk  \\
	\tnew \\
	\tmethod \\
	\tabsmethod \\
	\tintf \\
	\tanno
\end{mathpar}
\saveSpaceFig
\caption{Typing and subtyping of \name{}.}
\label{fig:typingrules}
\end{figure*}

@@

Generally, one can understand
$\mbody(m, I, J)$ in a way that it finds the most specific body of
method $m$, when the receiver has dynamic type $I$ and static type
$J$.  Notice, for example in \textsc{(S-Invk)}, that when
``$\new{I}$'' replaces ``\lstinline|this|'', its static type should be
the interface to which $m$ is dispatched. Therefore we also keep the
interface type in the definition of \mbody.

On the other hand, there is a relationship between path invocation and
the regular method invocation, and it can be observed from the
similarity between their semantic rules in
Figure~\ref{fig:smallstep}. For any \lstinline|e.I?m()|, the result
of evaluation remains unchanged if we set the static type of
\lstinline|e| to be $I$. This can be done by an implicit cast, that
is, we can define a function with one parameter type $I$, then
\lstinline|e| is passed to that function and directly returned. This
is equivalent to writing explicit casts like \lstinline|((I) e).m()|
in languages like Java.

@@

\subsection{Properties}
Previously the definitions of our model are given, now we proceed to prove the type soundness of 
the model, which relates typing to computation. The type soundness states that, if an expression is 
well-typed, then after many reduction steps it must reduce to a value, with its annotation to be a subtype of the original expression type.
Following the Featherweight Java paper~\cite{Igarashi01FJ}, the type-soundness theorem 
(Theorem~\ref{theorem_soundness}) is proved by using the standard technique of subject reduction (Theorem~\ref{theorem_subject})
and progress (Theorem~\ref{theorem_progress})~\cite{Wright1994}. In Theorem~\ref{theorem_progress} ``$\#(\overline{x})$'' denotes the length of
$\overline{x}$.
